%======================================================================
\chapter{Methods}
\label{ch:methods}
%======================================================================

%----------------------------------------------------------------------
\section{Spatial Discretization}

The bioheat PDE is discretized on a uniform Cartesian grid with
$N_x \times N_y$ voxels and spacing $\Delta x = L_x / N_x$,
$\Delta y = L_y / N_y$.
Interior voxels use 5-point standard second-order finite differences;
boundary voxels are modified by ghost-cell correction to enforce
the selected boundary condition (Section~\ref{sec:bc}).

The sparse system matrix $A_d = M^{-1}(K_d - W_b)$ is assembled once
at solver initialization and reused every timestep
(\texttt{physics/discretization.py}, \texttt{physics/bioheat.py}).

%----------------------------------------------------------------------
\section{Time Integration}

The discretized state equation~\eqref{eq:bioheat-discrete} is an ODE
of dimension $2N$, integrated forward in time with a fixed step $\Delta t$.
Two integrators are supported:

\begin{description}
    \item[Forward Euler] $\mathbf{x}_{k+1} = \mathbf{x}_k + \Delta t\,f(\mathbf{x}_k, P_k)$
        — first-order accurate, conditionally stable.
    \item[RK4 (default)] Standard 4th-order Runge-Kutta with
        zero-order hold on $P$ over each interval $[t_k,\, t_{k+1}]$.
\end{description}

The \gls{cfl} stability condition for the heat equation is:
\begin{equation}
    \Delta t \leq \frac{\rho c}{2k}
    \left(\frac{1}{\Delta x^{-2} + \Delta y^{-2}}\right)
\end{equation}
and is checked at startup (\texttt{utils/validators.py}).

%----------------------------------------------------------------------
\section{Solver Modes}

\subsection{Mode 0 — Open-Loop Bang-Bang}

A fixed power schedule (full power $P_{\max}$ for half the horizon,
then off) is applied without optimization.
Used as a baseline and for verifying physical correctness of the state
equations.

\subsection{Mode 1 — Indirect (PMP Gradient Projection)}

A forward-adjoint sweep algorithm iterates:
\begin{enumerate}
    \item \textbf{Forward pass:} integrate state equations~\eqref{eq:bioheat-discrete}
          and~\eqref{eq:arrhenius} under current $P(t)$.
    \item \textbf{Terminal costate:} evaluate $\boldsymbol{\lambda}_\Omega(t_f)$
          from the terminal cost gradient.
    \item \textbf{Backward pass:} integrate adjoint
          equations~\eqref{eq:adjoint-T}--\eqref{eq:adjoint-Omega} backward.
    \item \textbf{Control update:} project onto $[P_{\min}, P_{\max}]$ via~\eqref{eq:pontryagin}
          and update with step size $\beta$:
          \[P^{(k+1)} = (1-\beta)\,P^{(k)} + \beta\,P^*\]
    \item Repeat until $\|P^{(k+1)} - P^{(k)}\|_\infty < \varepsilon$.
\end{enumerate}
Convergence is sensitive to $\beta$ (default 0.15).
Because the Pontryagin control is near bang-bang, a large $\beta$ makes
the fixed-point iteration limit-cycle between ``under-ablated\,$\to$\,full
power'' and ``ablated\,$\to$\,off''; $\beta \approx 0.15$ damps this.
Oscillation indicates $\beta$ should be reduced further.

\subsection{Mode 2 — Direct Single-Shooting NLP}

The power schedule $\mathbf{u} \in \mathbb{R}^{n_{\text{steps}}}$ is passed
directly to \texttt{scipy.optimize.minimize} with method \texttt{SLSQP}.
The objective calls \texttt{forward\_simulate()} at every function evaluation,
making this mode computationally expensive but straightforward to implement.
Box constraints $P_{\min} \leq u_k \leq P_{\max}$ are enforced via
\texttt{scipy} bounds.
Because \gls{slsqp} approximates the gradient by finite differences
($\approx n+1$ forward PDE rollouts per gradient), this mode does not
scale to the $50^3$-voxel 3-D grid (Chapter~\ref{ch:results}).

%----------------------------------------------------------------------
\section{Constraint Checking}

After each simulation, the \texttt{ConstraintChecker} module evaluates:
\begin{description}
    \item[Path constraints] Maximum healthy-tissue temperature at each timestep;
        control bounds and slew-rate satisfaction.
    \item[Terminal constraints] Fraction of tumor voxels ablated
        ($\Omega_d(t_f) \geq 1$); total treatment time vs.\ $T_{\max}$.
\end{description}

%----------------------------------------------------------------------
\section{Software Architecture}

The codebase follows a strict separation of concerns.
Table~\ref{tab:modules} summarizes the module structure.

\begin{table}[htbp]
\centering
\caption{Module map}
\label{tab:modules}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{lll}
\toprule
Layer & Module & Role \\
\midrule
Config   & \texttt{config.py}                    & Single source of truth \\
Physics  & \texttt{physics/mesh.py}              & Grid, region masks \\
Physics  & \texttt{physics/discretization.py}    & Sparse $K_d$, $W_b$ assembly \\
Physics  & \texttt{physics/bioheat.py}           & Bioheat integrator \\
Physics  & \texttt{physics/arrhenius.py}         & Damage accumulation \\
Physics  & \texttt{physics/sar\_model.py}        & SAR field, $\mathbf{b}_P$ \\
Physics  & \texttt{physics/boundary\_conditions.py} & Dirichlet/Neumann/Robin \\
Control  & \texttt{control/ocp.py}               & Bolza cost $J$ \\
Control  & \texttt{control/adjoint.py}           & Costate ODEs, $P^*(t)$ \\
Control  & \texttt{control/constraints.py}       & Constraint verification \\
Control  & \texttt{control/solver.py}            & Three solver backends \\
Utils    & \texttt{utils/validators.py}          & CFL check \\
Utils    & \texttt{utils/io\_utils.py}           & Result serialization \\
\bottomrule
\end{tabular}
\end{table}

All modules import parameters from the global singleton \texttt{cfg}
(\texttt{from config import cfg}); no magic numbers appear outside
\texttt{config.py}.
