%======================================================================
\chapter{Conclusion and Future Work}
\label{ch:conclusion}
%======================================================================

%----------------------------------------------------------------------
\section{Summary}

This project developed a modular PDE-constrained optimal control
simulation for MRI-guided microwave thermal ablation, supporting both
2-D slices and full 3-D volumes.
The Pennes bioheat equation and Arrhenius damage ODE are discretized
by finite differences and integrated with a shared forward simulator
used by all three solver backends (open-loop, indirect PMP, direct NLP),
under a free-final-time formulation that truncates each trajectory at the
optimal stopping instant $t_f^\star$.
A full sweep of $3$ solvers $\times\,2$ dimensions $\times\,3$ probe
geometries (Chapter~\ref{ch:results}) benchmarks the methods under
matched Dirichlet boundary conditions and cost weights.

The indirect solver derives costate equations analytically from the
Pontryagin Maximum Principle and updates the control schedule via
forward-adjoint sweep iteration; it scales to 3-D in under an hour.
The direct solver delegates optimization to \texttt{scipy.optimize}
and trades speed for implementation simplicity, but does not scale to the
full 3-D grid.

%----------------------------------------------------------------------
\section{On the Difficulty of Tuning the Cost Weights}
\label{sec:gain-tuning}

The single most time-consuming part of this project was tuning the five
cost weights $(\alpha_1, \alpha_2, \alpha_3, \gamma_1, \gamma_2)$. Several
structural reasons make this hard.

\paragraph{Competing objectives with no common scale.}
The weights trade off four physically incommensurable quantities ---
ablation completeness, deposited energy, treatment time, and healthy-tissue
overheating --- each with different units and magnitudes. There is no
natural normalization, so the weights must be hand-balanced until each term
contributes a comparable amount to $J$; a change in one weight silently
shifts the operating point of all the others.

\paragraph{The ablation penalty is dimension-dependent.}
The incomplete-ablation term is a \emph{spatial integral},
$\gamma_1 \int_{\Omega_T}(1-\Omega_d)_+^2\,d\mathbf{r} \approx \gamma_1 N_T\,dV$,
so its value scales with the voxel measure $dV$. The weight
$\gamma_1 = 10^6$ was chosen so this term is $\approx 190$ for the 2-D area
element $dA \approx \SI{e-6}{\square\meter}$ --- large enough to dominate
the $O(10^1$--$10^2)$ energy and time costs and force the optimizer to
ablate. In 3-D the volume element collapses to
$dV = (\SI{2.5}{\mm})^3 \approx \SI{1.6e-8}{\cubic\meter}$, shrinking the
un-ablated penalty to $\gamma_1 N_T\,dV \approx 2.1$ --- now far smaller
than the $\sim\!200$ of energy required to ablate the volume. Doing nothing
then costs only $J \approx \gamma_2 t_f + \alpha_3 t_f + 2.1 \approx 35$ and
becomes the global optimum: this is exactly why the converged $20^3$ direct
runs in Chapter~\ref{ch:results} ablate \emph{nothing}. The same weight set
is correct in 2-D and degenerate in 3-D, so $\gamma_1$ must be re-scaled by
$dA/dV \approx 60\times$ when the dimensionality (or grid resolution)
changes.

\paragraph{A narrow feasible window.}
Even within a fixed dimension the usable range is narrow. The energy weight
$\alpha_1$ must be small enough that the optimizer is still willing to spend
power, yet large enough to separate the optimizers from the full-power
open-loop baseline; empirically the ablating optimum survives only for
$\alpha_1 \lesssim 1.2\times10^{-3}$, beyond which ``do nothing'' again
wins. The optimizers thus live on a knife-edge between two trivial
solutions (full power and zero power).

\paragraph{Near bang-bang control amplifies sensitivity.}
Because the Pontryagin control law is close to bang-bang, small weight
changes can flip an entire stretch of the schedule between $P_{\max}$ and
$0$. This is also why the indirect solver needs a heavily damped relaxation
($\beta = 0.15$): a larger step makes the fixed-point iteration limit-cycle
between ``under-ablated $\to$ full power'' and ``ablated $\to$ off.'' Weight
tuning and solver stability are therefore coupled, not independent.

%----------------------------------------------------------------------
\section{Remaining Challenges}

\begin{enumerate}
    \item \textbf{Indirect convergence:}
        The gradient projection algorithm is not globally convergent;
        the relaxation parameter $\beta$ requires manual tuning.
        Line-search or conjugate-gradient extensions would improve robustness.

    \item \textbf{Soft safety constraint:}
        $T \leq T_{\text{safe}}$ in healthy tissue is enforced as a soft
        $\alpha_2$ penalty, so the optimizer is permitted to exceed it
        when the ablation reward dominates. Chapter~\ref{ch:results} shows
        this is exactly what happens for the 3-D line and dipole sources,
        which overheat healthy tissue. Promoting the limit to a hard path
        constraint (as the time-optimal direct mode already does) would
        close this gap.

    \item \textbf{Direct single-shooting in 3-D:}
        Dense single-shooting re-integrates the full $50^3$-voxel PDE per
        finite-difference gradient and did not converge within nine hours
        for the 3-D runs; it converges only after the grid is coarsened to
        $20^3$. Direct \emph{collocation} (sparse NLP via IPOPT/CasADi) or
        the adjoint method --- which scaled to 3-D in under an hour --- is
        required at full resolution.

    \item \textbf{Cost-weight calibration:}
        As detailed in Section~\ref{sec:gain-tuning}, the weights are
        sensitive and dimension-dependent; in particular $\gamma_1$ must be
        re-scaled by the voxel volume when moving from 2-D to 3-D.
\end{enumerate}

%----------------------------------------------------------------------
\section{Future Work}

\begin{itemize}
    \item Replace single-shooting direct method with direct collocation
          (e.g., CasADi/IPOPT) for better scalability.
    \item \textbf{Receding-horizon MPC.} Close the loop by re-solving the
          OCP over a short prediction horizon at each control step, using
          the live MRI thermometry field as the initial condition. This
          naturally handles model error and measurement feedback and is the
          essential ingredient for real-time MRI-guided control.
    \item \textbf{Reinforcement learning.} Train a policy (e.g.\
          actor--critic or policy-gradient) on the discretized
          bioheat--damage Markov decision process, learning a power schedule
          directly from
          simulated rollouts. This avoids analytical gradients and is suited
          to handling imaging noise and patient-to-patient variability.
    \item Incorporate real MRI thermometry data (proton resonance
          frequency shift maps) as the feedback signal.
    \item Extend to heterogeneous tissue properties and patient-specific
          geometry from MRI segmentation.
    \item Implement safety-critical control (control barrier functions)
          to enforce the healthy-tissue temperature constraint as a
          hard constraint rather than a soft penalty.
    \item Re-scale the terminal ablation weight $\gamma_1$ by voxel volume
          so the cost functional is dimensionally consistent in 3-D.
\end{itemize}
