%======================================================================
\chapter{Introduction}
%======================================================================

%----------------------------------------------------------------------
\section{Motivation}

Liver cancer is among the leading causes of cancer-related mortality
worldwide~\cite{}.
Minimally invasive thermal ablation—delivering focused energy to destroy
tumor tissue without open surgery—is a standard treatment for unresectable
hepatocellular carcinoma.
Microwave ablation (MWA) deposits energy through dielectric heating and
produces predictable, roughly ellipsoidal ablation zones on the order of
\SIrange{2}{4}{\cm} in diameter.

\gls{mri} guidance is attractive because it provides superior soft-tissue
contrast, three-dimensional volumetric imaging, and real-time temperature
mapping via \emph{MR thermometry} (proton resonance frequency shift).
This combination—MWA plus MRI thermometry—creates a closed-loop opportunity:
the power schedule can be updated in real time based on measured temperature,
enabling precise ablation of the tumor while sparing healthy tissue.

%----------------------------------------------------------------------
\section{Problem Statement}

Given the temperature field measured by MRI thermometry, we seek a power
schedule $P(t) \in [0, P_{\max}]$ that:
\begin{enumerate}
    \item achieves complete thermal ablation of the tumor
          ($\Omega_d(\mathbf{r}, t_f) \geq 1$ for all $\mathbf{r} \in \Omega_T$);
    \item keeps healthy tissue below a safety temperature
          ($T(\mathbf{r},t) \leq T_{\text{safe}}$ for all $\mathbf{r} \in \Omega_H$);
    \item minimizes treatment duration and total energy deposition.
\end{enumerate}
This is a \gls{pde}-constrained \gls{ocp} with free final time,
state and control constraints, and a Bolza cost structure.

%----------------------------------------------------------------------
\section{Related Work}

Bioheat equation-based treatment planning dates to the work of
Pennes~\cite{} and the thermal dose formulation of
Sapareto and Dewey~\cite{}.
More recently, model-based optimal control approaches have been applied
to radiofrequency ablation~\cite{} and laser interstitial therapy~\cite{}.
Adjoint-based gradient methods for PDE-constrained control are surveyed
in~\cite{}.
MPC for thermal ablation using real-time MRI feedback was explored
in~\cite{}.
The present work integrates these threads into a unified simulation
framework with four solver modes (open-loop, indirect PMP, direct NLP, MPC)
and benchmarks them under matched conditions.

%----------------------------------------------------------------------
\section{Contributions}

The main contributions of this project are:
\begin{enumerate}
    \item A 2-D finite-difference simulator of the Pennes bioheat PDE
          and Arrhenius damage ODE, parameterized by a single configuration
          object with no magic numbers.
    \item An adjoint (costate) solver derived from the \gls{pmp} and
          verified against finite-difference gradient checks.
    \item Four interchangeable solver backends—open-loop, indirect gradient
          projection, direct single-shooting NLP, and receding-horizon
          \gls{mpc}—sharing a common result interface.
    \item Configurable boundary conditions (Dirichlet / Neumann / Robin)
          selectable at the command line.
    \item Quantitative comparison of the four solvers on a standard
          5~cm $\times$ 5~cm phantom with an 8~mm tumor.
\end{enumerate}

%----------------------------------------------------------------------
\section{Report Organization}

Chapter~\ref{ch:problem} states the mathematical problem.
Chapter~\ref{ch:methods} describes the numerical methods and software
architecture.
Chapter~\ref{ch:results} presents simulation results.
Chapter~\ref{ch:conclusion} concludes and outlines future work.
